\boldmath Disconnected pseudo-$C_\ell$ covariances for projected large-scale structure data

### 2019-06-23 draft/paper.tex

\documentclass[a4paper,11pt]{article}

\pdfoutput=1
\usepackage{jcappub}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amssymb,amsmath,bm,dsfont}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{balance}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage[caption=false]{subfig}

@@

\newcommand{\todo}[1]{{\bf TODO: #1}}
\newcommand{\red}[1]{{\color{red} #1}}

@@

@@

      \subsubsection{Covariances for flat skies}
        Similar results hold in the case of flat skies. As mentioned in Section \ref{ssec:theory.pcl}, in this case averaging over the Fourier-space azimuth happens while binning into bandpowers. Under the assumption that the underlying power spectra are roughly constant within each bandpower, in this case the covariance matrix of the bandpowers defined in Eq. \ref{eq:bandpowers} takes the form:
        \begin{equation}\label{eq:cov_flat}
          {\rm Cov}\left(\tilde{C}^{a_\alpha b_\beta}_q,\tilde{C}^{f_\phi g_\gamma}_{q'}\right) = C^{a_{\alpha'} f_{\phi'}}_{(q}C^{b_{\beta'} g_{\gamma'}}_{q')}\sum_{{\bf l}\in q} \frac{1}{N_q} \sum_{{\bf l}' \in q'} \frac{1}{N_{q'}}\,\,^{af}_{bg}\bar{\mathcal{W}}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}({\bf l},{\bf l}')+\left((g,\gamma)\leftrightarrow(f,\phi)\right),
        \end{equation}
        where the coefficients $\bar{\mathcal{W}}$ are related to the mode-coupling coefficients $\bar{\Xi}$ defined in \ref{eq:coeff_mcm_flat} in the same way that the curved-sky coefficients $\mathcal{W}$ were related to the $\Xi$:
        \begin{align}
          &\bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\delta  \delta  ,\delta  \delta  }=
           \bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\delta  \gamma_X,\delta  \gamma_X}=
           \bar{\mathcal{W}}^{\delta  \gamma_X,\delta  \gamma_X}_{\delta  \gamma_Y,\delta  \gamma_Y}=\bar{\mathcal{J}}^{00}\\
          &\bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=
           \bar{\mathcal{W}}^{\delta  \gamma_Z,\delta  \gamma_Z}_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=\bar{\mathcal{J}}^{0+}\\
          &\bar{\mathcal{W}}^{\gamma_X\gamma_Y,\gamma_X\gamma_Y}_{\gamma_W\gamma_Z,\gamma_W\gamma_Z}=\bar{\mathcal{J}}^{++}\\\nonumber
          &\bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
          &\hspace{12pt}
           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=-\bar{\mathcal{J}}^{--}\\\nonumber
          &\bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
          &\hspace{12pt}
           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=\bar{\mathcal{J}}^{--},
        \end{align}
        and, as before
        \begin{align}
          &\,^{af}_{bg}\bar{\mathcal{J}}^{00}_{{\bf l}{\bf l}'}=\bar{\Xi}^{00}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
           \,^{af}_{bg}\bar{\mathcal{J}}^{0+}_{{\bf l}{\bf l}'}=\bar{\Xi}^{0+}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\\
          &\,^{af}_{bg}\bar{\mathcal{J}}^{++}_{{\bf l}{\bf l}'}=\bar{\Xi}^{++}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
           \,^{af}_{bg}\bar{\mathcal{J}}^{--}_{{\bf l}{\bf l}'}=\bar{\Xi}^{--}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\\
          &\,^{af}_{bg}\bar{\mathcal{J}}^{+-}_{{\bf l}{\bf l}'}=
           \,^{af}_{bg}\bar{\mathcal{J}}^{-+}_{{\bf l}{\bf l}'}=
           \,^{af}_{bg}\bar{\mathcal{J}}^{0-}_{{\bf l}{\bf l}'}=0,
        \end{align}

@@

  \section{Discussion}\label{sec:discussion}
    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \todo{cites}, but it is particularly relevant for large-scale structure datasets. This problem is further complicated in this case, in comparison with e.g. CMB experiments, by two factors: the fact that the fields involved are non-Gaussian at some level and the higher complexity of the sky masks used in optical datasets. The main impact of survey geometry is the statistical coupling it induces between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable estimates of the posterior distribution for cosmological parameters. In this paper we have focused on the impact of survey geometry on the dominant Gaussian (i.e. disconnected) part of the covariance matrix.

@@

  \appendix
  \section{Flat sky}\label{app:flat}
    \begin{figure}
      \centering
      \includegraphics[width=0.6\textwidth]{./figures/mask-lss_flat1.pdf}
      \caption{Flat-sky mask used in the study of the analytical
        methods presented in this work in the flat-sky regime. The mask was
        constructed from the bright-object mask distributed with the first
        data release of the HSC collaboration \todo{cite} for the VVDS
        field. As in the full-sky case, it is used both for galaxy clustering
        and shear for simplicity.} \label{fig:mask_flat}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields. This figure additionally shows the $\chi^2$ distributions for power spectra involving $B$-modes. For the smaller scales covered by these flat-sky simulations, the NKA and spin-0 estimators work remarkably well, even for $B$-mode spectra.}
      \label{fig:chi2_1bin_flat}
    \end{figure}
    We have repeated the analysis described in Section \ref{sec:results} on flat-sky realizations, making use of the flat-sky implementation of {\tt NaMaster}. We generate Gaussian realizations of the galaxy overdensity and shear maps making use of flat-sky extensions of the methods described in Section \ref{ssec:results.sims}. In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \todo{cite} for the VVDS field, which is shown in Fig. \ref{fig:mask_flat}.

@@

  \section{Software implementation}\label{app:namaster}
    \todo{complete}

>>>>

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
\usepackage{jcappub}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{amssymb,amsmath,bm,dsfont}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{balance}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage[caption=false]{subfig}

@@

@@

        \begin{align}\label{eq:coeff_cmcm}

@@

      \subsubsection{Covariances for flat skies}
        Similar results hold in the case of flat skies. As mentioned in Section \ref{ssec:theory.pcl}, in this case averaging over the Fourier-space azimuth happens while binning into bandpowers. Under the assumption that the underlying power spectra are roughly constant within each bandpower, in this case the covariance matrix of the bandpowers defined in Eq. \ref{eq:bandpowers} takes the form:
        \begin{equation}\label{eq:cov_flat}
          {\rm Cov}\left(\tilde{C}^{a_\alpha b_\beta}_q,\tilde{C}^{f_\phi g_\gamma}_{q'}\right) = C^{a_{\alpha'} f_{\phi'}}_{(q}C^{b_{\beta'} g_{\gamma'}}_{q')}\sum_{{\bf l}\in q} \frac{1}{N_q} \sum_{{\bf l}' \in q'} \frac{1}{N_{q'}}\,\,^{af}_{bg}\bar{\mathcal{W}}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}({\bf l},{\bf l}')+\left((g,\gamma)\leftrightarrow(f,\phi)\right),
        \end{equation}
        where the coefficients $\bar{\mathcal{W}}$ are related to the mode-coupling coefficients $\bar{\Xi}$ defined in \ref{eq:coeff_mcm_flat} in the same way that the curved-sky coefficients $\mathcal{W}$ were related to the $\Xi$:
        \begin{align}
          &\bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\delta  \delta  ,\delta  \delta  }=
           \bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\delta  \gamma_X,\delta  \gamma_X}=
           \bar{\mathcal{W}}^{\delta  \gamma_X,\delta  \gamma_X}_{\delta  \gamma_Y,\delta  \gamma_Y}=\bar{\mathcal{J}}^{00}\\
          &\bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=
           \bar{\mathcal{W}}^{\delta  \gamma_Z,\delta  \gamma_Z}_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=\bar{\mathcal{J}}^{0+}\\
          &\bar{\mathcal{W}}^{\gamma_X\gamma_Y,\gamma_X\gamma_Y}_{\gamma_W\gamma_Z,\gamma_W\gamma_Z}=\bar{\mathcal{J}}^{++}\\\nonumber
          &\bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
          &\hspace{12pt}
           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=-\bar{\mathcal{J}}^{--}\\\nonumber
          &\bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
          &\hspace{12pt}
           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
           \bar{\mathcal{W}}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=\bar{\mathcal{J}}^{--},
        \end{align}
        and, as before
        \begin{align}\label{eq:coeff_cmcm_flat}
          &\,^{af}_{bg}\bar{\mathcal{J}}^{00}_{{\bf l}{\bf l}'}=\bar{\Xi}^{00}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
           \,^{af}_{bg}\bar{\mathcal{J}}^{0+}_{{\bf l}{\bf l}'}=\bar{\Xi}^{0+}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\\
          &\,^{af}_{bg}\bar{\mathcal{J}}^{++}_{{\bf l}{\bf l}'}=\bar{\Xi}^{++}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
           \,^{af}_{bg}\bar{\mathcal{J}}^{--}_{{\bf l}{\bf l}'}=\bar{\Xi}^{--}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\\
          &\,^{af}_{bg}\bar{\mathcal{J}}^{+-}_{{\bf l}{\bf l}'}=
           \,^{af}_{bg}\bar{\mathcal{J}}^{-+}_{{\bf l}{\bf l}'}=
           \,^{af}_{bg}\bar{\mathcal{J}}^{0-}_{{\bf l}{\bf l}'}=0,
        \end{align}

@@

  \section{Discussion}\label{sec:discussion}
    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \cite{2013PhRvD..88f3537D,2013MNRAS.432.1928T,2017arXiv170609359K}, but it is particularly relevant for large-scale structure datasets. This problem is further complicated in this case, in comparison with e.g. CMB experiments, by two factors: the fact that the fields involved are non-Gaussian at some level and the higher complexity of the sky masks used in optical datasets. The main impact of survey geometry is the statistical coupling it induces between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable estimates of the posterior distribution for cosmological parameters. In this paper we have focused on the impact of survey geometry on the dominant Gaussian (i.e. disconnected) part of the covariance matrix.

@@

  \appendix
  \section{Flat sky}\label{app:flat}
    \begin{figure}
      \centering
      \includegraphics[width=0.6\textwidth]{./figures/mask-lss_flat1.pdf}
      \caption{Flat-sky mask used in the study of the analytical methods presented in this work in the flat-sky regime. The mask was constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field. As in the full-sky case, it is used both for galaxy clustering and shear for simplicity.} \label{fig:mask_flat}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields. This figure additionally shows the $\chi^2$ distributions for power spectra involving $B$-modes. For the smaller scales covered by these flat-sky simulations, the NKA and spin-0 estimators work remarkably well, even for $B$-mode spectra.}
      \label{fig:chi2_1bin_flat}
    \end{figure}
    We have repeated the analysis described in Section \ref{sec:results} on flat-sky realizations, making use of the flat-sky implementation of {\tt NaMaster}. We generate Gaussian realizations of the galaxy overdensity and shear maps making use of flat-sky extensions of the methods described in Section \ref{ssec:results.sims}. In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field, which is shown in Fig. \ref{fig:mask_flat}.

@@

  \section{Software implementation}\label{app:namaster}
    In addition to the code functionality described in Section 3 of \cite{2019MNRAS.484.4127A}, we have now included the capability to estimate Gaussian covariance matrices using the NKA method. This functionality is structured around a {\tt python} class called {\tt NmtCovarianceWorkspace}. These objects are used to compute and store the covariance mode-coupling coefficients in Eqs. \ref{eq:coeff_cmcm} and \ref{eq:coeff_cmcm_flat}. They are initialized from two pairs of fields corresponding to the two power spectra for which the covariance is required. Once initialized, these coefficients can be reused for any other set of fields with the same combination of sky masks. {\tt NaMaster} then provides routines to estimate covariance matrix elements making use of the coupling coefficients stored in a {\tt NmtCovarianceWorkspace} object and best-guess power spectra for the fields involved using Eqs. \ref{eq:cov_sph} and \ref{eq:cov_flat}. Further details about the implementation and practical examples can be found in \url{https://namaster.readthedocs.io/en/latest/sample_covariance.html}.
